---
title: "peergraded"
author: "Rohan"
date: "24/10/2020"
output: html_document
---

### Introduction
This document has been submitted as part of the prediction assigment for Coursera's [Practical Machine Learning](https://www.coursera.org/learn/practical-machine-learning).

---

### Dataset background
> Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 
> One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
> Class A corresponds to the specified execution of the exercise, while the other 4 classes correspond to common mistakes.
The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.
---

### Assignment instructions
> One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants.
> The goal of your project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. You may use any of the other variables to predict with. You should create a report describing how you built your model, how you used cross validation, what you think the expected out of sample error is, and why you made the choices you did. You will also use your prediction model to predict 20 different test cases.
---


```{r}
library(caret)
library(knitr)

library(data.table)
library(rpart.plot)
library(rpart)

library(gbm)
library(ggplot2)

library(corrplot)

```
Now we will take the data and do the cleaning and then exploring the data. 

```{r}
testUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
traUrl  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"

test_this_data <- read.csv(url(testUrl))
tra_this_data <- read.csv(url(traUrl))


```
now proceeding for the cleaning the input of the data

```{r}
this_is_training_data <- tra_this_data[, colSums(is.na(tra_this_data)) == 0]
this_is_testing_data <- test_this_data[, colSums(is.na(test_this_data)) == 0]
```
now we will prepare the data for pred. in which we will consider seventy percentage of the data for the training set and rest of the thirty percentage of the data for the testing data set
and this_is_testing_data will be used furthuer again for the prediction of the 20 of the cases 

```{r}
this_is_training_data <- this_is_training_data[, -c(1:7)]
this_is_testing_data <- this_is_testing_data[, -c(1:7)]
dim(this_is_training_data)
```

```{r}
set.seed(1234)
datatraining <- createDataPartition(tra_this_data$classe, p = 0.7, list = FALSE)
this_is_training_data <- this_is_training_data[datatraining, ]
this_is_testing_data <- this_is_training_data[-datatraining, ]
dim(this_is_training_data)
dim(this_is_testing_data)
```
now we will be removing the variables that are non zero from the data gives
```{r}
noneZero <- nearZeroVar(this_is_training_data)
this_is_training_data <- this_is_training_data[, -noneZero]
this_is_testing_data <- this_is_testing_data[, -noneZero]
dim(this_is_training_data)
dim(this_is_testing_data)
```


```{r}
plot_cor <- cor(this_is_training_data[, -53])
corrplot(plot_cor, order = "FPC", method = "color", type = "upper", tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```

now in this as we can see that the corr. predic. are the ones with the dark colour intersec.

now we will be proceeding for the model building and for this we will use 2 different types of algorithms , trees and random forests for the prediction part 

```{r}
set.seed(20000)
tredec <- rpart(classe ~ ., data=this_is_training_data, method = "class")
rpart.plot(tredec)
```

now we will be validate the model 
```{r}
modelpre <- predict(tredec, this_is_testing_data, type = "class")
ab <- confusionMatrix(modelpre, this_is_testing_data$classe)
ab
```

```{r}
plot(modelpre)
```

now for the last part we will apply two models one by one 
the first one will be general boosted model and then the second one will be gbm model for this 
```{r}
set.seed(10000)
ctr_gbm <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
valid_gbm <- train(classe ~ .,data=this_is_training_data, method = "gbm", trControl = ctr_gbm, verbose = FALSE)
valid_gbm$finalModel
```
